Digging Deeper Into the State of the Practice for Domain Specific Research Software

### 2022-02-18 StateOfPractice/Papers/ICCS-2022-SOP_MethodologyPlusExamples/DigDeep.tex
Revisions to abstract

	We have developed a methodology for assessing the state of the software
	development practices for a given research software domain. Our methodology
	prescribes the following steps: i) Identify the domain; ii) Identify a list
	of candidate software packages; iii) Filter the list to a length of about 30
	packages; iv) Collect repository related data on each software package, like
	number of stars, number of open issues, number of lines of code; v) Fill in
	the measurement template (the template consists of 108 questions to assess 9
	qualities (including the qualities of installability, usability and
	visibility)); vi) Rank the software using the Analytic Hierarchy Process
	(AHP); vii) Interview developers (the interview consists of 20 questions and
	takes about an hour); and, viii) conduct a domain analysis. The collected
	data is analyzed by: i) comparing the ranking by best practices against an
	the ranking by popularity; ii) comparing artifacts, tools and processes to
	current research software development guidelines; and, iii) exploring pain
	points. We estimate the time to complete an assessment for a given domain at
	173 person hours. The method is illustrated via the example of Lattice
	Boltzmann Solvers, where we find that the top packages engaged in most of
	recommended best practices, but still show room for improvement with respect
	to providing API documentation, a roadmap, a code of conduct, programming
	style guide, uninstall instructions and continuous integration.

>>>>

	We have developed a methodology for assessing the state of the software
	development practice for a given research software domain. Our methodology
	prescribes the following steps: i) Identify the domain; ii) Identify a list
	of candidate software packages; iii) Filter the list to a length of about 30
	packages; iv) Collect repository related data on each software package, like
	number of stars, number of open issues, number of lines of code; v) Fill in
	the measurement template (the template consists of 108 questions to assess 9
	qualities (including the qualities of installability, usability and
	visibility)); vi) Rank the software using the Analytic Hierarchy Process
	(AHP); vii) Interview developers (the interview consists of 20 questions and
	takes about an hour); and, viii) Conduct a domain analysis. The collected
	data is analyzed by: i) comparing the ranking by best practices against 
	the ranking by popularity; ii) comparing artifacts, tools and processes to
	current research software development guidelines; and, iii) exploring pain
	points. We estimate the time to complete an assessment for a given domain at
	173 person hours. The method is illustrated via the example of Lattice
	Boltzmann Solvers, where we find that the top packages engaged in most of
	recommended best practices, but still show room for improvement with respect
	to providing API documentation, a roadmap, a code of conduct, programming
	style guide, uninstall instructions and continuous integration.
